% T I T L E   P A G E
% -------------------
% Last updated August 24, 2023, by IST-Client Services
% The title page is counted as page `i' but we need to suppress the
% page number. Also, we don't want any headers or footers.
\pagestyle{empty}
\pagenumbering{roman}

% The contents of the title page are specified in the "titlepage"
% environment.
\begin{titlepage}
        \begin{center}
        \vspace*{1.0cm}

        \Huge
        {\bf Detecting Unchecked Exception-Related Behavioural Breaking Changes with UnCheckGuard }

        \vspace*{1.0cm}

        \normalsize
        by \\

        \vspace*{1.0cm}

        \Large
        Vinayak Sharma \\

        \vspace*{3.0cm}

        \normalsize
        A thesis \\
        presented to the University of Waterloo \\ 
        in fulfillment of the \\
        thesis requirement for the degree of \\
        Master of Applied Science \\
        in \\
        Electrical and Computer Engineering \\

        \vspace*{1.0cm}

        Waterloo, Ontario, Canada, 2025 \\

        \vspace*{1.0cm}

        \copyright\ Vinayak Sharma 2025 \\
        \end{center}
\end{titlepage}

% The rest of the front pages should contain no headers and be numbered using Roman numerals starting with `ii'
\pagestyle{plain}
\setcounter{page}{2}

\cleardoublepage % Ends the current page and causes all figures and tables that have so far appeared in the input to be printed.
% In a two-sided printing style, it also makes the next page a right-hand (odd-numbered) page, producing a blank page if necessary.
\phantomsection    % allows hyperref to link to the correct page
 

% D E C L A R A T I O N   P A G E
% -------------------------------
  % The following is a sample Declaration Page as provided by the GSPA
  % December 13th, 2006.  It is designed for an electronic thesis.
 \addcontentsline{toc}{chapter}{Author's Declaration}
 \begin{center}\textbf{Author's Declaration}\end{center}

 % Author's Declaration Option ONE - line 118:  
 \noindent
This thesis consists of material all of which I authored or co-authored: see Statement of
Contributions included in the thesis. This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.
  % Author's Declaration Option TWO - line 121. Updated August 21st, 2023. Use the following declaration text if appropriate by removing the percent character and space at the beginning of line 121, and add a percent symbol and space at line 118 to change Author's Declaration Option ONE to a remark that is not printed.
 \noindent  
% This thesis consists of material all of which I authored or co-authored: see Statement of Contributions included in the thesis. This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.
  \bigskip
  
  \noindent
I understand that my thesis may be made electronically available to the public.

\cleardoublepage
\phantomsection    % allows hyperref to link to the correct page

% S T A T E M E N T  O F  C O N T R I B U T I O N S
% -------------------------------
  % The following is a sample Declaration Page as provided by the GSPA
  % December 13th, 2006.  It is designed for an electronic thesis.
 \addcontentsline{toc}{chapter}{Statement of Contributions}
 \begin{center}\textbf{Statement of Contributions}\end{center}

This thesis primarily comprises chapters derived from our research paper which has been
accepted at the 25th IEEE International Conference on Source Code Analysis and Manipulation (SCAM 2025), incorporating wording changes, stylistic updates, and other
modifications.

I have added the following additional work:

\begin{itemize}
  \item Added information about software libraries and an overview of the thesis in Chapter~\ref{intro}.
  \item Introduced Chapter~\ref{background} to explain the core concepts related to the thesis.
  \item Elaborated further on the motivating example in Chapter~\ref{example}.
  \item Included examples and explanations on data collection in Chapter~\ref{data}.
  \item Described the problem definition and setup information related to taint analysis in Chapter~\ref{methodology}.
  \item Added one more related work in Chapter~\ref{related}
\end{itemize}

\cleardoublepage
\phantomsection    % allows hyperref to link to the correct page

% A B S T R A C T
% ---------------
\addcontentsline{toc}{chapter}{Abstract}
\begin{center}\textbf{Abstract}\end{center}

      The ubiquitous use of third-party libraries in software development has enabled developers to quickly add
      new functionality to their client software. Unfortunately, library usage also carries a cost in
      terms of software maintenance: library upgrades may include breaking changes, in which client expectations
      about library behaviour are no longer met in new library versions. Behavioural breaking
      changes can be particularly insidious, and in their full generality, could require sophisticated program
      analysis techniques to (approximately) detect.

      In this work, we present our UnCheckGuard tool, which detects a class of behavioural breaking changes---those
      related to exceptions thrown by Java libraries. UnCheckGuard analyzes both sides of the library/client
      duet. On the library side, UnCheckGuard creates a list of new exceptions that may be thrown by methods
      in a library's public API, including by its transitive callees. On the client side, UnCheckGuard identifies
      client methods that call library methods with new exceptions. To reduce false positives, UnCheckGuard
      additionally filters out new exceptions that cannot be triggered by particular clients, using taint analysis. It therefore can be
      used by client developers as a tool to screen library updates for relevant incompatibilities.

      We have evaluated UnCheckGuard on 302 libraries and 352 library-client pairs drawn from the DUETS collection
      and found 120 libraries with newly-added exceptions, as well as 1708 callsites to library methods which,
      when upgraded to the latest version, may introduce
      a behavioural breaking change in the client due to a newly added unchecked exception. These findings
      highlight the practical value of UnCheckGuard in identifying exception-related incompatibilities
      introduced by library upgrades.

\cleardoublepage
\phantomsection    % allows hyperref to link to the correct page

% A C K N O W L E D G E M E N T S
% -------------------------------
\addcontentsline{toc}{chapter}{Acknowledgements}
\begin{center}\textbf{Acknowledgements}\end{center}

First and foremost, I would like to thank Dr. Patrick Lam for his unwavering support throughout every stage of my master's journey. His constant positivity and encouragement toward excellence made it possible for me to work on such an interesting topic. From narrowing down the research focus to perfecting the tool, he has guided me at every step of the research process. His mentorship has not only shaped the direction of my thesis but has also played a significant role in my growth as a researcher and academic.

I would also like to thank my parents and my sister for their continuous support. As professors, my parents have always inspired me to strive for academic excellence and have been a constant source of motivation throughout my life. My sister, in particular, has provided me with valuable advice on how to improve academically, which has been instrumental during this journey.

Lastly, I would like to thank Dr. Weiyi Shang and Dr. Derek Rayside for their time and thoughtful feedback in reviewing my thesis.
\cleardoublepage
\phantomsection    % allows hyperref to link to the correct page


% D E D I C A T I O N
% -------------------------------
\addcontentsline{toc}{chapter}{Dedication}
\begin{center}\textbf{Dedication}\end{center}
This thesis is dedicated to my family and Addy.
\cleardoublepage
\phantomsection    % allows hyperref to link to the correct page



% T A B L E   O F   C O N T E N T S
% ---------------------------------
\renewcommand\contentsname{Table of Contents}
\tableofcontents
\cleardoublepage
\phantomsection    % allows hyperref to link to the correct page

% L I S T   O F   F I G U R E S
% -----------------------------
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page

% L I S T   O F   T A B L E S
% ---------------------------
\addcontentsline{toc}{chapter}{List of Tables}
\listoftables
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page

% L I S T   O F   A B B R E V I A T I O N S
% ---------------------------
\renewcommand*{\abbreviationsname}{List of Abbreviations}
\printglossary[type=abbreviations]
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page

% L I S T   O F   S Y M B O L S
% ---------------------------
\printglossary[type=symbols]
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page


% Change page numbering back to Arabic numerals
\pagenumbering{arabic}

